---
title: "Project 2"
author: "Nicholas Ilvento"
date: '2022-12-06'
output:
  pdf_document: default
  html_document: default
---

```{r}
library(mosaic)
library(DescTools)
library(binom)
library(boot)
library(TeachingDemos)
library(exactci)
library(PropCIs)
library(XNomial)
library(EMT)
library(readxl)
library(SciViews)
library(epitools)
library(perm)
library(combinat)
library(purrr)
library(MonteCarlo)
library(survival)
library(survminer)
library(car)
library(MASS)
```

```{r}
telco=read_excel("C:/Users/nicki/OneDrive/Documents/School/R/MA-460/Telco.xlsx")
telco
```

```{r}
favstats(telco$age)
favstats(telco$income)
favstats(telco$employ)
```

```{r}
mean_vec=c(mean(telco$age),mean(telco$income),mean(telco$employ))
sd_vec=c(sd(telco$age),sd(telco$income),sd(telco$employ))
median_vec=c(median(telco$age),median(telco$income),median(telco$employ))
max_vec=c(max(telco$age),max(telco$income),max(telco$employ))
min_vec=c(min(telco$age),min(telco$income),min(telco$employ))
stat_data=matrix(data=c(mean_vec,sd_vec,median_vec,max_vec,min_vec),nrow=3,ncol=5)
colnames(stat_data)=c("mean","sd","median","max","min")
rownames(stat_data)=c("age","income","employ")
stat_data
```

```{r}
age_churn=c()
income_churn=c()
employ_churn=c()
for(i in 1:length(telco$age)){
  if(telco$churn[i]=="1"){
    age_churn=c(age_churn,telco$age[i])
    income_churn=c(income_churn,telco$income[i])
    employ_churn=c(employ_churn,telco$employ[i])
  }
}
mean_vec=c(mean(age_churn),mean(income_churn),mean(employ_churn))
sd_vec=c(sd(age_churn),sd(income_churn),sd(employ_churn))
median_vec=c(median(age_churn),median(income_churn),median(employ_churn))
max_vec=c(max(age_churn),max(income_churn),max(employ_churn))
min_vec=c(min(age_churn),min(income_churn),min(employ_churn))
stat_data_churn=matrix(data=c(mean_vec,sd_vec,median_vec,max_vec,min_vec),nrow=3,ncol=5)
colnames(stat_data_churn)=c("mean","sd","median","max","min")
rownames(stat_data_churn)=c("age","income","employ")
stat_data_churn
```
```{r}
count(telco$churn=="1")
```

```{r}
cor(telco$age,telco$income)
cor(telco$age,telco$employ)
cor(telco$income,telco$employ)
```

```{r}
box_plot_data=data.frame(income=telco$income,age=telco$age,employ=telco$employ)
```

```{r}
boxplot(box_plot_data$age,box_plot_data$employ)
```

```{r}
count(telco$marital==0) #unmarried?
count(telco$marital==1) #married?
```

```{r}
count(telco$gender==0) #male?
count(telco$gender==1) #female?
```

```{r}
hist(telco$marital,xlab="Marital Status",ylab="Count",main="Histogram of Marital Status",breaks=c(0,0.5,1))
hist(telco$gender,xlab="Gender",ylab="Count",main="Histogram of Gender",breaks=c(0,0.5,1))
```

```{r}
telco$marital_cat=c(rep(0,length(telco$marital)))
telco$gender_cat=c(rep(0,length(telco$gender)))
for(i in 1:length(telco$marital)){
  if(telco$marital[i]==0){
    telco$marital_cat[i]="unmarried"
  }
  else(telco$marital_cat[i]="married")
  if(telco$gender[i]==0){
    telco$gender_cat[i]="male"
  }
  else(telco$gender_cat[i]="female")
}
```

```{r}
Gender=telco$gender
Marital_Status=telco$marital
ggplot(data.frame(telco),aes(x=Marital_Status))+geom_bar()
ggplot(data.frame(telco),aes(x=Gender))+geom_bar()
```

```{r}
gender_churn=table(telco$churn,telco$gender)
gender_churn_total=table(telco$churn,telco$gender)
gender_churn_total=rbind(gender_churn_total,c(colSums(gender_churn_total)))
gender_churn_total=cbind(gender_churn_total,c(rowSums(gender_churn_total)))
rownames(gender_churn_total)=c("no churn","churn","TOTAL")
colnames(gender_churn_total)=c("male","female","TOTAL")
gender_churn_total
```

```{r}
marital_churn=table(telco$churn,telco$marital)
marital_churn_total=table(telco$churn,telco$marital)
marital_churn_total=rbind(marital_churn_total,c(colSums(marital_churn_total)))
marital_churn_total=cbind(marital_churn_total,c(rowSums(marital_churn_total)))
rownames(marital_churn_total)=c("no churn","churn","TOTAL")
colnames(marital_churn_total)=c("unmarried","married","TOTAL")
marital_churn_total
```

##patterns
Females churn more. Unmarried people churn more. 125/500 = 1/4 of people churn.

##statistically significant association
```{r}
chisq.test(x=telco$gender,y=telco$churn)
chisq.test(x=telco$marital,y=telco$churn)
```

```{r}
chisq.test(gender_churn)
chisq.test(marital_churn)
```

```{r}
gender_marital=table(telco$marital,telco$gender)
gender_marital_total=table(telco$marital,telco$gender)
gender_marital_total=rbind(gender_marital_total,c(colSums(gender_marital_total)))
gender_marital_total=cbind(gender_marital_total,c(rowSums(gender_marital_total)))
rownames(gender_marital_total)=c("unmarried","married","TOTAL")
colnames(gender_marital_total)=c("male","female","TOTAL")
gender_marital_total
chisq.test(gender_marital)
```

```{r}
unmarried_female_churn=0
unmarried_male_churn=0
married_female_churn=0
married_male_churn=0
for(i in 1:length(telco$churn)){
  if(telco$churn[i]==1&&telco$gender[i]==1&&telco$marital[i]==0){
    unmarried_female_churn=unmarried_female_churn+1
  }
  if(telco$churn[i]==1&&telco$gender[i]==0&&telco$marital[i]==0){
    unmarried_male_churn=unmarried_male_churn+1
  }
  if(telco$churn[i]==1&&telco$gender[i]==1&&telco$marital[i]==1){
    married_female_churn=married_female_churn+1
  }
  if(telco$churn[i]==1&&telco$gender[i]==0&&telco$marital[i]==1){
    married_male_churn=married_male_churn+1
  }
}
unmarried_female_churn
unmarried_male_churn
married_female_churn
married_male_churn
```

```{r}
model=survfit(Surv(tenure,churn)~1,data=telco)
plot(model,lty=1:4,xlab="Time",ylab="Survival Prob",ylim=c(0,1))
```

```{r}
model=survfit(Surv(tenure,churn)~gender,data=telco)
plot(model,lty=1:4,xlab="Time",ylab="Survival Prob",ylim=c(0,1))
legend(50,0.95,legend=c("male","female"),lty=1:2,cex=0.8)
```

```{r}
survdiff(Surv(tenure,churn)~gender,data=telco)
```

Ho: The different levels are proportional
Ha: The different levels are not proportional
```{r}
pairwise_survdiff(Surv(tenure,churn)~gender,data=telco,p.adjust.method="bonferroni")
```
With a p-value of 0.75, we do not have enough evidence to reject the claim that the two levels are proportional with each other.

```{r}
model=survfit(Surv(tenure,churn)~marital,data=telco)
plot(model,lty=1:4,xlab="Time",ylab="Survival Prob",ylim=c(0,1))
legend(55,0.95,legend=c("unmarried","married"),lty=1:2,cex=0.8)
```

```{r}
survdiff(Surv(tenure,churn)~marital,data=telco)
```

Ho: The different levels are proportional
Ha: The different levels are not proportional
```{r}
pairwise_survdiff(Surv(tenure,churn)~marital,data=telco,p.adjust.method="bonferroni")
```
With a p-value of 0.043, we have enough evidence to reject the null of proportional hazards and say that the two levels are not proportional with each other.















